---
title: "Business Intelligence Lab Submission Markdown"
author: "<Specify your group name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** 
|                                                   |                                                                                                                                                                          |
|                                                   | 1.  134780 - C - Trevor Okinda                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 2.  132840 - C - Sheila Wangui                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 3.  131749 - C - Teresia Nungari                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 4.  135203 - C - Tom Arnold                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   |                                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   | Lakers                                                                                                       
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                                                                                                  |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                                                                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                                                                                               |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# Setup Chunk

We start by installing all the required packages

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## mlbench ----
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## readr ----
if (require("readr")) {
  require("readr")
} else {
  install.packages("readr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## e1071 ----
if (require("e1071")) {
  require("e1071")
} else {
  install.packages("e1071", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## factoextra ----
if (require("factoextra")) {
  require("factoextra")
} else {
  install.packages("factoextra", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## FactoMineR ----
if (require("FactoMineR")) {
  require("FactoMineR")
} else {
  install.packages("FactoMineR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```

------------------------------------------------------------------------

**Note:** the following "*R Markdown*" options have been set as the defaults in this markdown:

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

# Load dataset

Load the student performance dataset

```{r Load dataset}
library(readr)
library(dplyr)
student_performance_dataset <-
  readr::read_csv(
    "../data/StudentPerformanceDataset.csv", 
    col_types =
      readr::cols(
        class_group =
          readr::col_factor(levels = c("A", "B", "C")),
        gender = readr::col_factor(levels = c("1", "0")),
        YOB = readr::col_date(format = "%Y"),
        regret_choosing_bi =
          readr::col_factor(levels = c("1", "0")),
        drop_bi_now =
          readr::col_factor(levels = c("1", "0")),
        motivator =
          readr::col_factor(levels = c("1", "0")),
        read_content_before_lecture =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        anticipate_test_questions =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        answer_rhetorical_questions =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        find_terms_I_do_not_know =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        copy_new_terms_in_reading_notebook =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        take_quizzes_and_use_results =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        reorganise_course_outline =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        write_down_important_points =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        space_out_revision =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        studying_in_study_group =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        schedule_appointments =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        goal_oriented =
          readr::col_factor(levels =
                              c("1", "0")),
        spaced_repetition =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        testing_and_active_recall =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        interleaving =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        categorizing =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        retrospective_timetable =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        cornell_notes =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        sq3r = readr::col_factor(levels =
                                   c("1", "2", "3", "4")),
        commute = readr::col_factor(levels =
                                      c("1", "2",
                                        "3", "4")),
        study_time = readr::col_factor(levels =
                                         c("1", "2",
                                           "3", "4")),
        repeats_since_Y1 = readr::col_integer(),
        paid_tuition = readr::col_factor(levels =
                                           c("0", "1")),
        free_tuition = readr::col_factor(levels =
                                           c("0", "1")),
        extra_curricular = readr::col_factor(levels =
                                               c("0", "1")),
        sports_extra_curricular =
          readr::col_factor(levels = c("0", "1")),
        exercise_per_week = readr::col_factor(levels =
                                                c("0", "1",
                                                  "2",
                                                  "3")),
        meditate = readr::col_factor(levels =
                                       c("0", "1",
                                         "2", "3")),
        pray = readr::col_factor(levels =
                                   c("0", "1",
                                     "2", "3")),
        internet = readr::col_factor(levels =
                                       c("0", "1")),
        laptop = readr::col_factor(levels = c("0", "1")),
        family_relationships =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        friendships = readr::col_factor(levels =
                                          c("1", "2", "3",
                                            "4", "5")),
        romantic_relationships =
          readr::col_factor(levels =
                              c("0", "1", "2", "3", "4")),
        spiritual_wellnes =
          readr::col_factor(levels = c("1", "2", "3",
                                       "4", "5")),
        financial_wellness =
          readr::col_factor(levels = c("1", "2", "3",
                                       "4", "5")),
        health = readr::col_factor(levels = c("1", "2",
                                              "3", "4",
                                              "5")),
        day_out = readr::col_factor(levels = c("0", "1",
                                               "2", "3")),
        night_out = readr::col_factor(levels = c("0",
                                                 "1", "2",
                                                 "3")),
        alcohol_or_narcotics =
          readr::col_factor(levels = c("0", "1", "2", "3")),
        mentor = readr::col_factor(levels = c("0", "1")),
        mentor_meetings = readr::col_factor(levels =
                                              c("0", "1",
                                                "2", "3")),
        `Attendance Waiver Granted: 1 = Yes, 0 = No` =
          readr::col_factor(levels = c("0", "1")),
        GRADE = readr::col_factor(levels =
                                    c("A", "B", "C", "D",
                                      "E"))),
    locale = readr::locale())
```

## Scale Data Transform


```{r Scale Data Transform}
# BEFORE
summary(student_performance_dataset)

model_of_the_transform <- preProcess(student_performance_dataset, method = c("scale"))
print(model_of_the_transform)
student_perfomance_dataset_scale_transform <- predict(model_of_the_transform,
                                          student_performance_dataset)
# AFTER
summary(student_perfomance_dataset_scale_transform)

```

Next, Center Data Transform

```{r Center Data Transform}
# BEFORE
summary(student_performance_dataset)

model_of_the_transform <- preProcess(student_performance_dataset, method = c("center"))
print(model_of_the_transform)
student_performance_dataset_center_transform <- predict(model_of_the_transform, # nolint
                                           student_performance_dataset)

# AFTER
summary(student_performance_dataset_center_transform)

```

# Standardize Data Transform


```{r Standardize Data Transform}
# BEFORE
summary(student_performance_dataset)


model_of_the_transform <- preProcess(student_performance_dataset,
                                     method = c("scale", "center"))
print(model_of_the_transform)
student_performance_dataset_standardize_transform <- predict(model_of_the_transform, # nolint
                                                student_performance_dataset)
# AFTER
summary(student_performance_dataset_standardize_transform)

```

## Normalize Data Transform

```{r Normalize Data Transform}
#### The Normalize Transform on the Student Performance Dataset ----
summary(student_performance_dataset)
model_of_the_transform <- preProcess(student_performance_dataset, method = c("range"))
print(model_of_the_transform)
student_performance_dataset_normalize_transform <- predict(model_of_the_transform, # nolint
                                              student_performance_dataset)
summary(student_performance_dataset_normalize_transform)

```

## Box-Cox Power Transform

```{r Box-Cox Power Transform}
summary(student_performance_dataset)


model_of_the_transform <- preProcess(student_performance_dataset, method = c("BoxCox"))
print(model_of_the_transform)
student_performance_dataset_box_cox_transform <- predict(model_of_the_transform, # nolint
                                                         student_performance_dataset)

# AFTER
summary(student_performance_dataset_box_cox_transform)

```

**etc.** as per the lab submission requirements. Be neat and communicate in a clear and logical manner.