---
title: "Business Intelligence Project"
author: "<Specify your name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |                         |
|----------------------------------------------|-------------------------|
| **Student ID Number**                        | 134780                  |
| **Student Name**                             | Trevor Augustine Okinda |
| **BBIT 4.2 Group**                           | C                       |
| **BI Project Group Name/ID (if applicable)** | ...                     |

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults: <BR> `knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)
```

# Understanding the Dataset (Exploratory Data Analysis (EDA))

## Loading the Dataset

### Source:

The dataset that was used can be downloaded here: <https://www.kaggle.com/datasets/bhanupratapbiswas/weather-data?resource=download>

### Reference:

Biswas Bhanupratap (2023). Weather Data - Understanding and Utilizing Weather Data [Dataset]. Kaggle. <https://www.kaggle.com/datasets/bhanupratapbiswas/weather-data?resource=download> #no lint

```{r Dataset Loader}
library(readr)
Weather_Data <- read_csv("Weather_Data.csv", 
                         col_types = cols(Visibility_km = col_double(), 
                                          Weather = col_character()))
View(Weather_Data)
```

...to be continued
