---
title: "Business Intelligence Lab Submission Markdown"
author: "<Specify your group name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** 
|                                                   |                                                                                                                                                                          |
|                                                   | 1.  134780 - C - Trevor Okinda                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 2.  132840 - C - Sheila Wangui                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 3.  131749 - C - Teresia Nungari                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 4.  135203 - C - Tom Arnold                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   |                                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   | Lakers                                                                                                       
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                                                                                                  |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                                                                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                                                                                               |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# Setup Chunk

We start by installing all the required packages

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## mlbench ----
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## klaR ----
if (require("klaR")) {
  require("klaR")
} else {
  install.packages("klaR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## e1071 ----
if (require("e1071")) {
  require("e1071")
} else {
  install.packages("e1071", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## readr ----
if (require("readr")) {
  require("readr")
} else {
  install.packages("readr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## LiblineaR ----
if (require("LiblineaR")) {
  require("LiblineaR")
} else {
  install.packages("LiblineaR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## naivebayes ----
if (require("naivebayes")) {
  require("naivebayes")
} else {
  install.packages("naivebayes", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```

------------------------------------------------------------------------

**Note:** the following "*R Markdown*" options have been set as the defaults in this markdown:

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

# Load dataset

Load the PimaIndiansDiabetes dataset and check its structure

```{r Load dataset}
library(mlbench)
data("PimaIndiansDiabetes")
# Check the structure of the dataset
str(PimaIndiansDiabetes)
```

## Split dataset


```{r Split dataset}
# Split the dataset into training and testing sets
set.seed(123)  # Set a random seed for reproducibility


# Assuming "diabetes" is the target variable
inTrain <- createDataPartition(y = PimaIndiansDiabetes$diabetes, p = 0.6, list = FALSE)
training_data <- PimaIndiansDiabetes[inTrain, ]
testing_data <- PimaIndiansDiabetes[-inTrain, ]

```

Next, we create a Naive Bayes classifier using the Pima Indians Diabetes dataset

```{r Naive Bayes classifier}
# Create a Naive Bayes classifier using the Pima Indians Diabetes dataset
pima_nb_model <- e1071::naiveBayes(diabetes ~ pregnant + glucose + pressure + triceps + insulin + mass + pedigree + age, data = PimaIndiansDiabetes)

# Check the summary of the Naive Bayes model
summary(pima_nb_model)

# Assuming you have already trained the Naive Bayes model (pima_nb_model) and have a testing dataset (testing_data)
predictions_nb_caret <- predict(pima_nb_model, newdata = testing_data[, c("pregnant", "glucose", "pressure", "triceps", "insulin", "mass", "pedigree", "age")])


# Calculate the confusion matrix
confusion_matrix <- confusionMatrix(predictions_nb_caret, testing_data$diabetes)
print(confusion_matrix)

# Create a plot based on the confusion matrix
plot(confusion_matrix$table)

```

# Bootstrapping train control
Training a linear regression model (for regression)


```{r Bootstrapping train control}
train_control <- trainControl(method = "boot", number = 500)

pima_lm_model <- caret::train(diabetes ~
                                              pregnant + glucose + pressure + triceps + insulin +
                                              mass + pedigree + age,
                                            data = training_data,
                                            trControl = train_control,
                                            na.action = na.omit, method = "glm", metric = "Accuracy")
# Test the trained linear regression model using the testing dataset
predictions_lm <- predict(pima_lm_model, newdata = testing_data[, c("pregnant", "glucose", "pressure", "triceps", "insulin", "mass", "pedigree", "age")])


#View the predicted values for the observations
print(predictions_lm)

```

## Binary Classification: Logistic Regression with 10-fold cross-validation

```{r Normalize Data Transform}
train_control <- trainControl(method = "cv", number = 10, classProbs = TRUE, summaryFunction = twoClassSummary)

pima_lr_model <- caret::train(diabetes ~ .,
                              data = training_data,
                              trControl = train_control,
                              na.action = na.omit,
                              method = "glm", 
                              family = "binomial", 
                              metric = "ROC")

# Test the trained logistic regression model using the testing dataset
predictions_lr <- predict(pima_lr_model, newdata = testing_data[, c("pregnant", "glucose", "pressure", "triceps", "insulin", "mass", "pedigree", "age")])

# Print the model and predicted values
print(pima_lr_model)
print(predictions_lr)

```

## LDA classifier based on a 5-fold cross-validation

```{r Box-Cox Power Transform}
train_control <- trainControl(method = "cv", number = 5)

pima_lda_model <- caret::train(diabetes ~ ., data = training_data,
                               trControl = train_control, na.action = na.omit,
                               method = "lda", metric = "Accuracy")

# Test the trained LDA model using the testing dataset
predictions_lda <- predict(pima_lda_model, newdata = testing_data[, c("pregnant", "glucose", "pressure", "triceps", "insulin", "mass", "pedigree", "age")])

# View the summary of the model
print(pima_lda_model)

# View the confusion matrix
confusion_matrix <- confusionMatrix(predictions_lda, testing_data$diabetes)
print(confusion_matrix)

```

**etc.** as per the lab submission requirements. Be neat and communicate in a clear and logical manner.